---
title: "Data_transform_Rfunction"
output: html_document
date: "2023-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Function for dataframe transformation

We have created an R function [transform_data()] that takes a dataframe and a vector of column-names to be selected as inputs. It returns a dataset with the selected columns and all lowercase characters in the column names.

<center>
**transform_data(data, selected_cols)**
</center>

Note: Users would need to install and load the 'tidyverse' package to ensure the function works as intended.

### Let's test out the funtion
Let's use an in-built R dataset, called 'iris' to test the output of our function. First, let's view the iris datset before the transformation.
```{r cars}
head(iris)
```
Now, let's apply our function to the iris dataset. Before that, we must identify our inputs for the function.
For example, if we wish to get Sepal.Length, Sepal.Width, Species columns from the iris dataset, we can define our input variables as follows:
<center>
**data = iris**

**selected_cols = c('Sepal.Length', 'Sepal.Width', 'Species')**
</center>
```{r echo=TRUE}
#User would need to install and load the 'tidyverse' package to ensure the function works as intended.
#Comment out the next line, if you don't have the 'tidyverse' package installed
#install.packages("tidyverse")
library(tidyverse)

# The function takes a dataframe and a vector of column names to be selected as arguments 
# Returns dataset with selected columns and all lowercase characters in column names.
transform_data <- function(data, selected_cols) {
  if (is.data.frame(data)) {
    # Rename all columns to lowercase
    new_data <- data %>% select(selected_cols)
    colnames(new_data) <- tolower(colnames(new_data))
    return(new_data)
  } else {
    # Return the original data if it's not a data frame
    return(print("Input is not a data frame"))
  }
}

# Let's use the function to transform the column names
selectedcols <- c('Sepal.Length', 'Sepal.Width', 'Species')
new_iris <- transform_data(iris, selected_cols = selectedcols)

# Let's compare and validate the transformation.
head(new_iris)
```
The functions works! It returned the specified columns, with lowercase column names.
